Fractional Types

### 2016-12-19 typespp.tex
bibtex
Add dual of (#α)ₖ rule
tweak one sentence

Both groupoids involve some notion of ``division'' that can be captured at the
level of types using a syntactic notion of \emph{fractional types}. Our aim is
to explore this notion of types and its associated operational semantics. The
remainder of the paper is organized as follows. We start by reviewing the
necessary background consisting of the language $\Pi$ for programming with
isomorphisms or equivalences in a reversible information-preserving way. Sec.~4
explains the main novel semantic ideas of using $\Pi$ programs to generate
non-trivial groupoids with fractional cardinality. Sec.~5 translates the
semantic ideas into an extension of $\Pi$ with new type constructors denoting
non-trivial groupoids and new programs that manipulate such types. Sec.~6
presents the operational semantics of the extended language. The last two
sections put our work in perspective and conclude.

@@

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. We recall that, as all fractional types are inhabited (see Remark
in Sec.~\ref{sec:symmetryg}), and thus there is no loss in generality in going
to pointed types. We then allow universal quantification over indices $k$
associated to values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same~$k$ but are
otherwise arbitrary. Once we introduce a universal quantifier in one direction
of execution, we are led to also include an existential quantifier for the
reverse execution. Intuitively, a forward evaluation step that is at liberty to
choose a $k$ will, when reversed encounter a particular, i.e., existentially
quantified, $k$. The syntax is therefore as follows.

>>>>

Both groupoids involve some notion of ``division'' that can be captured at the
level of types using a syntactic notion of \emph{fractional types}. Our aim is
to explore this notion of types and its associated operational semantics. The
remainder of the paper is organized as follows. We start by reviewing the
necessary background consisting of the language $\Pi$ for programming with
isomorphisms or equivalences in a reversible information-preserving way. Sec.~4
explains the main novel semantic ideas of using $\Pi$ programs to generate
non-trivial groupoids with fractional cardinality. Sec.~5 translates the
semantic ideas into an extension of $\Pi$ with new type constructors denoting
non-trivial groupoids and new programs that manipulate such types. Sec.~6
presents the operational semantics of the extended language. The last section
puts our work in perspective and concludes.

@@

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. As all fractional types are inhabited (see Remark
in Sec.~\ref{sec:symmetryg}), there is no loss in generality in going
to pointed types. We then allow universal quantification over indices $k$
associated to values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same~$k$ but are
otherwise arbitrary. Once we introduce a universal quantifier in one direction
of execution, we are led to also include an existential quantifier for the
reverse execution. Intuitively, a forward evaluation step that is at liberty to
choose a $k$ will, when reversed encounter a particular, i.e., existentially
quantified, $k$. The syntax is therefore as follows.
